\documentclass[11pt]{article}

\usepackage[norsk]{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{graphicx, float}
\usepackage{pstricks-add}


\author{Kjetil Kjeka}
\title{TTK4130 - Exercise 2}
\date{\today}


%Slik at matriser kan defineres med linjer
\makeatletter
\renewcommand*\env@matrix[1][*\c@MaxMatrixCols c]{%
  \hskip -\arraycolsep
  \let\@ifnextchar\new@ifnextchar
  \array{#1}}
\makeatother




\begin{document}
\maketitle
\section*{Problem 1}
\subsection*{a}
There are 5 variables, so 5 equations should be implemented.

\section*{Problem 2}
For a rational transfer function to be positive real it must satisfy:
\begin{enumerate}
\item All the poles of $H(s)$ have real parts less than or equal to zero.
\item $Re(H(j\omega)) \geq 0$ for all $\omega$ so that $j \omega$ is not a pole of $H(s)$.
\item If $j \omega_0$ is pole in $H(s)$, then it is a simple pole, and
\[Res_{s = j\omega_0} [H(s)] = \lim_{s \rightarrow j\omega_0} (s - j\omega_0)H(s)\]
is real and positive. If $H(s)$ has a pole at inifinity, then it is a simple pole, and
\[R_{\infty} = \lim_{\omega \rightarrow \infty} \frac{H(j\omega)}{j\omega}\]
\end{enumerate}

\subsection*{a}
In regards to the transfer function:
\[H_1(s) = \frac{1}{1+Ts}\]
It is obvious that all poles have real parts less then or equal to zero as long as $T > 0$. The frequency response
\[H_1(j\omega) = \frac{1 - j \omega T}{1 + (\omega T)^2}\]
has positive real parts for all $\omega$
\[Re(H_1(j \omega)) = \frac{1}{1 + (\omega T)^2}\]
And it got no poles on the imaginary axis. Thus $H_1$ is real positive.

The other transfer function:
\[H_2(s) = \frac{s}{s^2 + \omega^2_0}\]
Has poles only on the imaginary axis, $j\omega$ and $-j\omega$. The first and second criteria is thus satisfied but the third criteria remains.
\[Res_{s = \pm j \omega_0}H(s) = \frac{1}{2}\]
Thus $H_2$ is real positive.

\subsection*{b}
Consider the transfer function
\[H_3(s) = \frac{s + a}{(s + b)(s + c)} \]
Since $c > 0$ and $b > 0$ the first criteria is satisfied and the thrid criteria is not applicable. Considering the second criteria:
\[Re(H_3(j\omega)) = \frac{abc - a\omega^2 + b\omega^2 + c\omega^2 }{(\omega^2 + b^2)(\omega^2 + c^2)} \]
If $a \leq b + c$ the the criteria is satisfied, and if $a > b + c$ it fails. Meaning it is positive real for all $a \leq b + c$

\subsection*{c}
Considering the transfer function
\[H_4(s) = \frac{s^2 + a^2}{s(s^2 + \omega_0^2)}\]
The first criteria is satisfied since all poles are on the imaginary axis. The second criteria is also satsified since $Re(H_4(j \omega)) = 0$. The third criteria requires that the following are real and positive:
\[Res_{s = 0}[H_4(s)] = \frac{a^2}{\omega_0^2}\]
\[Res_{s = \pm j\omega_0}[H_4(s)] = \frac{\omega_0^2 - a^2}{2\omega_0^2}\]
Thus if $a < \omega_0$ then the transfer function is positive real.

\subsection*{d}

By chosing the storage function
\[V = \frac{1}{2} T y^2\]
i get 
\[\dot{V} = Ty = uy - y^2\]
Which obviously satisfy
\[\dot{V} = uy - g(y)\]
where $g(y) = y^2$ which is positive definite.

\subsection*{e}
It's already given that all poles are in the left half plane thus the first criteria is satisfied and the third is not applicable. So wether $H(s)$ is positive real is the same as if the second criteria is satisfied, namely if $Re[H(j\omega)] \geq 0$ for all $\omega \not = 0$ 





\end{document}